\chapter*{Acknowledgements}

I would like to express my sincere gratitude to all the individuals who have supported and guided me throughout the journey of completing this undergraduate project. Their encouragement and assistance have been invaluable in shaping this work.

First and foremost, I am deeply grateful to my family for their unwavering love and constant support. A special thanks to my parents, Márcio and Martina, for their endless encouragement and inspiration, which have been the foundation of my success. I am thankful for their sacrifices, which have allowed me to pursue my academic endeavours wholeheartedly.

I am indebted to all my teachers and professors throughout my academic journey from Escola Tangará to the Federal University of Rio de Janeiro, passing by Escola Parque Ipiranga, Colégio Dom Bosco, pH, Pensi, and Poliedro. Their knowledge, passion, and dedication to teaching have played an instrumental role in shaping my intellectual and personal growth.

I would like to extend my appreciation to all my friends, who have been a source of inspiration, motivation and joy. Their friendship and humour have provided much-needed relief during life's challenging moments.

I am grateful to my Glance Team colleagues for all the shared knowledge and experience that modelled my professional growth. Working with them and sharing the happiest years of my life has been a great pleasure. Thanks to Bárbara Tenório, Gabriel Levcovitz, Gabriel Oliveira, Gabriela Lemos, Gustavo Machado, and Leandro Alves for being my family in Geneva. Gabriel Silva and Marcelo Teixeira also deserve special thanks for being present not only as collaborators in this work but as a great company while living at Saint-Genis-Pouilly.

I would like to express my deepest gratitude to Carmen Maidantchik for giving me an opportunity that changed my personal and professional life and for all the work done on the collaboration between UFRJ and CERN. I am also thankful CERN employees that I met during the five years working at Glance. Special thanks to Adriana Telesca for trusting me with the developer job at the ALICE experiment and for sharing her experience with me.

Thanks to my undergraduate project advisor, Flávio de Mello, for his guidance, patience and insightful feedback to ensure the success of this project.

Finally, my personal thanks go to Michelly Texeira for being part of this work from the idea, execution, and until the final words of this text. I am genuinely grateful for her presence in my life. 
